---
title: OSM Resource
author: Dr. Emily Burchfield
date: '2020-07-16'
slug: osm
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-07-16T20:46:45-04:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r echo=F, warning=F, message=F}
shp_list <- Sys.glob("C:/Users/eburchf/OneDrive - Emory University/WF/Bangladesh/data/*.shp")
```


```{r message=F, warning=F}
library(tidyverse)
library(sf)
```

I just discovered [Geofabrik's](https://www.geofabrik.de/) clean daily shapefiles of [OpenStreetMap data](https://www.openstreetmap.org/#map=7/23.544/90.344) for basically every country on EARTH.  In this post, I'll work with data from Bangladesh which you can download by clicking the `[.shp.zip]` file next to Bangladesh  [here](http://download.geofabrik.de/asia.html). As you start playing with this data, you'll probably also want to review the metadata for the shapefiles, which you can find [here](download.geofabrik.de/osm-data-in-gis-formats-free.pdf). I dropped the downloaded shapefiles in a `data` folder on my machine. Note that all files for Bangladesh are around 2.7 GB, so if you're following along on a storage-challenged machine, you might want to go with a smaller country.  

Let's start by inspecting what's in this folder.  We can print out all of the files with extension `.shp` by running the following code (remember, this assumes you've moved all of your data to a `/data` sub-folder in your working directory):

```{r eval=F}
shp_list <- Sys.glob("./data/*.shp")
```

```{r}
shp_list
```

Lots to dig through.  We can also use this new `my_shp` list to help us open files.  Let's build a function to help us inspect the contents of each shapefile.

```{r}
# Note that some of the shapefiles might take awhile to load
inspect_shp <- function(dir) {
  
  shp <- st_read(dir, quiet = T)
  print(paste0(dir, " is a ", st_geometry_type(shp)[1], " with ", nrow(shp), " observations."))
  
  fclass_list <- unique(shp$fclass)
  
  shp_plt <- ggplot(shp) +
    geom_sf() +
    theme_minimal() +
    labs(title = dir,
         subtitle = "From Geofabrik.de")

  # dump anything you want to return from the function in this list
  out_list <- list(shp, fclass_list, shp_plt)
  
  return(out_list)
}
```

Let's try out our `inspect_shp()` function:

```{r}
out <- inspect_shp(shp_list[2])
```

```{r}
out[[2]]
```

```{r}
out[[3]]
```

Let's inspect some of the other datasets:

```{r message=F, warning=F}
places <- inspect_shp(shp_list[5])
places[[2]]
places[[3]]
```

```{r message=F, warning=F}
water <- inspect_shp(shp_list[17])
water[[2]]
water[[3]]
```

```{r message=F, warning=F}
roads <- inspect_shp(shp_list[12])
roads[[2]]
roads[[3]]
```

If you want to learn more about working with OSM in `R`, you can also check out the [`osmdata`](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html) package, which looks really helpful if you want to pull data directly from OSM.

