---
title: OSM Resource
author: Dr. Emily Burchfield
date: '2020-07-16'
slug: osm
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-07-16T20:46:45-04:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r echo=F}
setwd("C:/Users/eburchf/OneDrive - Emory University/WF/Bangladesh/")
```


```{r message=F, warning=F}
library(tidyverse)
library(sf)

roi <- st_read("./data/roi.shp", quiet=T)
```

I just discovered Geofabrik's clean daily shapefiles of [OpenStreetMap data](https://www.openstreetmap.org/#map=7/23.544/90.344) for basically every country on EARTH.  In this post, I'll work with data from Bangladesh which you can download by clicking the `[.shp.zip]` file next to Bangladesh  [here](http://download.geofabrik.de/asia.html). As you start playing with this data, you'll probably also want to review the metadata for the shapefiles, which you can find [here](download.geofabrik.de/osm-data-in-gis-formats-free.pdf). I dropped the downloaded shapefiles in a `data` folder on my machine. Note that all files for Bangladesh are around 2.7 GB, so if you're following along on a storage-challenged machine, you might want to go with a smaller country.  Note that I'm also using a region of interest (ROI) shapefile to trim down the county-wide shapefiles.  If you have a ROI of interest, just replace my ROI shapefile directory path with your ROI shapefile directory path.  Be sure to check projection!

Let's start by inspecting what's in this folder.  We can print out all of the files with extension `.shp` by running the following code (remember, this assumes you've moved all of your data to a `/data` sub-folder in your working directory):

```{r}
shp_list <- Sys.glob("./data/*.shp")
shp_list
```
Lots to dig through.  We can also use this new `my_shp` list to help us open files.  Let's build a function to help us inspect the contents of each shapefile.

```{r}
# Note that some of the shapefiles might take awhile to load
inspect_shp <- function(dir) {
  
  shp <- st_read(dir, quiet = T)
  print(paste0(dir, " is a ", st_geometry_type(shp)[1], " with ", nrow(shp), " observations."))
  
  fclass_list <- unique(shp$fclass)
  
  shp_plt <- ggplot(shp) +
    geom_sf() +
    theme_minimal() +
    labs(title = dir,
         subtitle = "From Geofabrik.de")

  # dump anything you want to return from the function in this list
  out_list <- list(shp, fclass_list, shp_plt)
  
  return(out_list)
}
```

Let's try out our `inspect_shp()` function:

```{r}
out <- inspect_shp(shp_list[2])
```

```{r}
out[[2]]
```

```{r}
out[[3]]
```

We can also use functions to clean up the shapefiles in a systematic way, e.g. reproject or crop.

```{r}
clean_shp <- function(shp, roi) {
  # reproject
  extent <- st_transform(roi, st_crs(shp))
  # and crop
  shp_c <- st_crop(shp, extent)
  # I won't do this here, but you could add any other tidying you needed to do here
  return(shp_c)
}
```

Let's clean and visualize the landuse data for my ROI:

```{r message=F, warning=F}
lulc <- clean_shp(out[[1]], roi) 

ggplot() +
  geom_sf(data = lulc, aes(fill = fclass)) +
  theme_minimal() +
  labs(title = "Land use",
       subtitle = "Geofrabrik download on 7/16/20",
       fill = "")
```

This is a quick way to confirm that coverage ain't great in my ROI.  Let's inspect some of the other datasets:

```{r message=F, warning=F}
places <- inspect_shp(shp_list[5])
places[[2]]
places[[3]]
```
```{r message=F, warning=F}
water <- inspect_shp(shp_list[17])
water[[2]]
water[[3]]
```

```{r message=F, warning=F}
roads <- inspect_shp(shp_list[12])
roads[[2]]
roads[[3]]
```

If you want to learn more about working with OSM in `R`, you can also check out the [`osmdata`](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html) package, which looks really helpful if you want to pull data directly from OSM.

